\Large Taking uncertainty seriously \newline \normalsize A Bayesian approach to word embedding bias estimation

### 2021-08-05 presentationESSLLI21/presentationESSLLI.Rmd
Adds up-to-date version

---
title: "\\Large Taking uncertainty seriously  \\newline \\normalsize A  Bayesian
  approach to  word embedding bias estimation "
author: "Alicja Dobrzeniecka \\& Rafal  Urbaniak \\footnotesize \\newline (LoPSE research group, University of  Gdansk)" 
date:  "ExpSem2021, ESSLLI"
fontsize: 10pt
output:  
  beamer_presentation:
    theme: Rafal_beamerSly1
    keep_tex: yes
    slide_level: 2
classoption: x11names, dvipsnames, bibspacing,natbib
urlcolor: blue
bibliography: [../references/cosineReferences1.bib]
csl: ../references/apa-6th-edition.csl 
link-citations: yes
---

@@

- Introduction to the topic of bias in word embeddings
- Analysis of methodological problems with MAC implementation
- Explanation of pre-averaging limitations in methods such as WEAT and MAC
- Our research and Bayesian approach

@@

<!-- <!-- (the gender direction is the top principal compontent of ten gender pair difference vectors).  --> -->

<!-- -  Given the (ideally) gender neutral words $N$ and the gender direction $g$ the direct gender bias is: -->
<!-- <!-- ($c$ is a parameter determining how strict we want to be): --> -->

@@

##   Cosine distance and  connection type

## Cosine distance and  connection type

@@

- Introduced to the topic of bias in word embeddings
- Analysed methodological problems with MAC implementation
- Explained of pre-averaging limitations in methods such as WEAT and MAC
- Proposed Bayesian approach

>>>>

---
title: '\Large Taking uncertainty seriously  \newline \normalsize A  Bayesian approach
  to  word embedding bias estimation '
author: Alicja Dobrzeniecka \& Rafal  Urbaniak \footnotesize \newline (LoPSE research
  group, University of  Gdansk)
date: "ExpSem2021, ESSLLI"
output:
  beamer_presentation:
    theme: Rafal_beamerSly1
    keep_tex: yes
    slide_level: 2
  slidy_presentation: default
fontsize: 10pt
classoption: x11names, dvipsnames, bibspacing,natbib
urlcolor: blue
bibliography: ../references/cosineReferences1.bib
csl: ../references/apa-6th-edition.csl
link-citations: yes
---

@@

- Introduction to the topic of bias in word embeddings
- WEAT and MAC methods
- Methodological problems with MAC implementation
- Limitations of pre-averaging in bias detection methods
- Including uncertainty with Bayesian approach

@@

<!-- <!-- (the gender direction is the top principal compontent of ten gender pair difference vectors).  -->

<!-- -  Given the (ideally) gender neutral words $N$ and the gender direction $g$ the direct gender bias is: -->
<!-- <!-- ($c$ is a parameter determining how strict we want to be): -->

@@

@@

- Introduction to the topic of bias in word embeddings
- WEAT and MAC methods
- Methodological problems with MAC implementation
- Limitations of pre-averaging in bias detection methods
- Including uncertainty with Bayesian approach

### 2021-08-05 presentationESSLLI21/presentationESSLLI.tex
Adds up-to-date version

\begin{frame}{The plan of the presentation}
\protect\hypertarget{the-plan-of-the-presentation}{}
\begin{itemize}
\tightlist
\item
  Introduction to the topic of bias in word embeddings
\item
  Analysis of methodological problems with MAC implementation
\item
  Explanation of pre-averaging limitations in methods such as WEAT and
  MAC
\item
  Our research and Bayesian approach
\end{itemize}
\end{frame}

@@

\begin{center}\includegraphics[width=0.6\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-1-1} \end{center}
\normalsize

--\textgreater{}

--\textgreater{}
\end{block}
\end{frame}

@@

\begin{frame}{Some methodological problems}
\protect\hypertarget{some-methodological-problems-2}{}
\begin{block}{Outliers and surprisingly dissimilar words}
\protect\hypertarget{outliers-and-surprisingly-dissimilar-words}{}
We study those by visualizations and uncertainty estimates

\begin{block}{No principled interpretation}
\protect\hypertarget{no-principled-interpretation}{}
\begin{longtable}[]{@{}ll@{}}
\toprule
Religion Debiasing & MAC (distance) \\
\midrule
\endhead
Biased & 0.859 \\
Hard Debiased & 0.934 \\
Soft Debiased (\(\lambda\) = 0.2) & 0.894 \\
\bottomrule
\end{longtable}

@@

\begin{frame}{The problem with pre-averaging}
\protect\hypertarget{the-problem-with-pre-averaging}{}
\begin{block}{Key conceptual issues}
\protect\hypertarget{key-conceptual-issues}{}
\begin{itemize}
\tightlist
\item
  It throws away information about sample sizes
\item
  It ignores variation in the raw data, which leads to false confidence
\end{itemize}
\end{block}

@@

\begin{center}\includegraphics[width=0.7\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-4-1} \end{center}
\normalsize

@@

\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-6-1} \end{center}
\normalsize
\end{block}
\end{frame}

\begin{frame}{Cosine distance and connection type}
\protect\hypertarget{cosine-distance-and-connection-type}{}
\begin{block}{Distances for ``muslim''}
\protect\hypertarget{distances-for-muslim}{}
\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-7-1} \end{center}
\end{block}
\end{frame}

\begin{frame}{Cosine distance and connection type}
\protect\hypertarget{cosine-distance-and-connection-type-1}{}
\begin{block}{Distances for ``priest''}
\protect\hypertarget{distances-for-priest}{}
\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-8-1} \end{center}
\end{block}
\end{frame}

>>>>

\begin{frame}{The plan of the presentation}
\protect\hypertarget{the-plan-of-the-presentation}{}
\begin{itemize}
\tightlist
\item
  Introduction to the topic of bias in word embeddings
\item
  WEAT and MAC methods
\item
  Methodological problems with MAC implementation
\item
  Limitations of pre-averaging in biad detection methods
\item
  Bayesian approach
\end{itemize}
\end{frame}

@@

\begin{center}\includegraphics[width=0.6\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-1-1} \end{center}
\normalsize
\end{block}
\end{frame}

@@

\begin{frame}{Some methodological problems}
\protect\hypertarget{some-methodological-problems-2}{}
\begin{block}{Outliers and surprisingly dissimilar words}
\protect\hypertarget{outliers-and-surprisingly-dissimilar-words}{}
We study those by visualizations and uncertainty estimates
\end{block}
\end{frame}

\begin{frame}{Some methodological problems}
\protect\hypertarget{some-methodological-problems-3}{}
\begin{block}{Distances for ``muslim''}
\protect\hypertarget{distances-for-muslim}{}
\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-4-1} \end{center}
\end{block}
\end{frame}

\begin{frame}{Some methodological problems}
\protect\hypertarget{some-methodological-problems-4}{}
\begin{block}{Distances for ``priest''}
\protect\hypertarget{distances-for-priest}{}
\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-5-1} \end{center}
\end{block}
\end{frame}

\begin{frame}{Some methodological problems}
\protect\hypertarget{some-methodological-problems-5}{}
\begin{block}{No principled interpretation}
\protect\hypertarget{no-principled-interpretation}{}
\begin{longtable}[]{@{}ll@{}}
\toprule
Religion Debiasing & MAC (distance) \\
\midrule
\endhead
Biased & 0.859 \\
Hard Debiased & 0.934 \\
Soft Debiased (\(\lambda\) = 0.2) & 0.894 \\
\bottomrule
\end{longtable}

@@

\begin{frame}{The problem with pre-averaging}
\protect\hypertarget{the-problem-with-pre-averaging}{}
\begin{block}{Key conceptual issues}
\protect\hypertarget{key-conceptual-issues}{}
\begin{itemize}
\tightlist
\item
  It throws away information about sample sizes
\item
  It ignores variation in the raw data, which leads to false confidence
\end{itemize}

@@

\begin{center}\includegraphics[width=0.7\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-6-1} \end{center}
\normalsize

@@

\begin{center}\includegraphics[width=0.8\linewidth]{presentationESSLLI_files/figure-beamer/unnamed-chunk-8-1} \end{center}
\normalsize
\end{block}
\end{frame}
